---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data
Loading the raw data.
```{r}
rawData <- read.csv("activity.csv", stringsAsFactors = FALSE)
```
Printing out some information about `rawData`.
```{r}
summary(rawData)
```
Checking the type of features in `rawData`.
```{r  collapse = TRUE}
class(rawData$steps)
class(rawData$date)
class(rawData$interval)
```
Transforming the `rawData$date` feature from a `character` format to a `Date` format.
```{r collapse=TRUE}
rawData$date <- as.Date(rawData$date)
class(rawData$date)
```

## What is mean total number of steps taken per day?
Calculating the total number of steps taken per day (ignoring missing values)
```{r}
stepsPerDay  <- aggregate(steps ~ date, data=rawData, sum, na.action = na.omit)
head(stepsPerDay)
```
Make an histogram of the total number of steps taken each day.
```{r histogram, fig.height=4}
hist(stepsPerDay$steps, breaks=50, main="Total number of steps taken each day", xlab = "steps")
```  

```{r echo = FALSE}
stepsPerDayMean <- mean(stepsPerDay$steps)
stepsPerDayMedian <- median(stepsPerDay$steps)
```
Calculate and report the mean and median of the total number of steps taken per day. 
```{r collapse=TRUE}
mean(stepsPerDay$steps)
median(stepsPerDay$steps)
```

## What is the average daily activity pattern?
Calculating the average number of steps taken per interval - averaged across all days (ignoring missing values).
```{r}
stepsPerInterval  <- aggregate(steps ~ interval, data=rawData, mean, na.action = na.omit)
head(stepsPerInterval)
```
The time series plot  of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis) is shown below.
```{r timeseries, fig.height= 4}
plot(stepsPerInterval$interval, stepsPerInterval$steps, type = "l", main="Average number of steps per 5-minute interval", xlab="5-minute interval", ylab="average no of steps")
```

```{r echo=FALSE}
max_interval <- stepsPerInterval[which.max(stepsPerInterval$steps),1]
max_steps <- stepsPerInterval[which.max(stepsPerInterval$steps),2]
```
Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps? It is the interval `r max_interval` with `r max_steps` steps.

```{r collapse=TRUE}
stepsPerInterval[which.max(stepsPerInterval$steps),]
```


## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
